---
title: "P8105 - Homework 6"
author: "Joe LaRocca"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)

```

## Problem 1

### Upload Weather Data

```{r}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```

### Create Function to Return Log(b_0 * b_1)

```{r}

bootstrap_log = function(lm) {
  log_b0_b1 = lm |>
    broom::tidy() |>
    pull(estimate) |>
    prod() |>
    log()
  return(log_b0_b1)
}

```

### Create Function to Return R^2 value

```{r}

bootstrap_rsq = function(lm) {
  r_squared = lm |>
    broom::glance() |>
    pull(r.squared)
  return(r_squared)
}

```

### Bootstrap 5000 Samples

```{r}

bootstraps_weather = weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin, data = df)),
    log_b0_b1 = map(models, bootstrap_log),
    r_sq = map(models, bootstrap_rsq)
  ) |>
  select(-strap, -models) |>
  unnest(log_b0_b1) |>
  unnest(r_sq)
  
```

### Plot the Distribution of log_b0_b1

```{r}

bootstraps_weather |>
  ggplot(aes(x = log_b0_b1)) + 
  geom_histogram(fill = "red", col = "black") + 
  labs(
    title = "Distribution of Log(b0 + b1) in a Bootstrapped Sample of 5000 SLRs",
    x = "Log(b0 * b1)",
    y = "Count"
  )

```

### Plot the Distribution of r_sq

```{r}

bootstraps_weather |>
  ggplot(aes(x = r_sq)) + 
  geom_histogram(fill = "lightblue", col = "black") + 
    labs(
    title = "Distribution of R^2 in a Bootstrapped Sample of 5000 SLRs",
    x = "R^2",
    y = "Count"
  )

```

We can see that both of these plots look to have an approximately Normal distribution, i.e. the distributions of $log(b_0 * b_1)$ and $R^2$ are approximately Normal in our sample of 5000 linear models.

### Identifying Quantiles for Confidence Intervals

```{r}

bootstraps_weather |>
  summarize(
    ci_lower_log_b0_b1 = quantile(log_b0_b1, 0.025),
    ci_upper_log_b0_b1 = quantile(log_b0_b1, 0.975),
    ci_lower_r_sq = quantile(r_sq, 0.025),
    ci_upper_r_sq = quantile(r_sq, 0.975)
  ) |>
  knitr::kable(digits = 3)

```

We can see that the 95\% confidence interval for $log(b_0 * b_1)$ is (1.967, 2.058), while the 95\% confidence interval for $R^2$ is (0.894, 0.927).

## Problem 2



## Problem 3
