---
title: "P8105 - Homework 6"
author: "Joe LaRocca"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)
```

## Problem 1

### Upload Weather Data

```{r}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```

### Create Function to Return Log(b_0 * b_1)

```{r}

bootstrap_log = function(lm) {
  log_b0_b1 = lm |>
    broom::tidy() |>
    pull(estimate) |>
    prod() |>
    log()
  return(log_b0_b1)
}

```

### Create Function to Return R^2 value

```{r}

bootstrap_rsq = function(lm) {
  r_squared = lm |>
    broom::glance() |>
    pull(r.squared)
  return(r_squared)
}

```

### Bootstrap 5000 Samples

```{r}

bootstraps_weather = weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin, data = df)),
    log_b0_b1 = map(models, bootstrap_log),
    r_sq = map(models, bootstrap_rsq)
  ) |>
  select(-strap, -models) |>
  unnest(log_b0_b1) |>
  unnest(r_sq)
  
```

### Plot the Distribution of log_b0_b1

```{r}

bootstraps_weather |>
  ggplot(aes(x = log_b0_b1)) + 
  geom_histogram(fill = "red", col = "black") + 
  labs(
    title = "Distribution of Log(b0 + b1) in a Bootstrapped Sample of 5000 SLRs",
    x = "Log(b0 * b1)",
    y = "Count"
  )

```

### Plot the Distribution of r_sq

```{r}

bootstraps_weather |>
  ggplot(aes(x = r_sq)) + 
  geom_histogram(fill = "lightblue", col = "black") + 
    labs(
    title = "Distribution of R^2 in a Bootstrapped Sample of 5000 SLRs",
    x = "R^2",
    y = "Count"
  )

```

We can see that both of these plots look to have an approximately Normal distribution, i.e. the distributions of $log(b_0 * b_1)$ and $R^2$ are approximately Normal in our sample of 5000 linear models.

### Identifying Quantiles for Confidence Intervals

```{r}

bootstraps_weather |>
  summarize(
    ci_lower_log_b0_b1 = quantile(log_b0_b1, 0.025),
    ci_upper_log_b0_b1 = quantile(log_b0_b1, 0.975),
    ci_lower_r_sq = quantile(r_sq, 0.025),
    ci_upper_r_sq = quantile(r_sq, 0.975)
  ) |>
  knitr::kable(digits = 3)

```

We can see that the 95\% confidence interval for $log(b_0 * b_1)$ is (1.967, 2.058), while the 95\% confidence interval for $R^2$ is (0.894, 0.927).

## Problem 2

### Upload and Filter Data

```{r}

homicides = read_csv("data/homicide-data.csv") |> 
  unite(col = "city_state", city, state, sep = ", ") |>
  filter(
    city_state != "Dallas, TX",
    city_state != "Phoenix, AZ",
    city_state != "Kansas City, MO",
    city_state != "Tulsa, AL",
    victim_race == "White" | victim_race == "Black",
    victim_sex == "Male" | victim_sex == "Female"
  ) |> 
  mutate(victim_age = as.numeric(victim_age)) |>
  mutate(unsolved = 
      disposition == "Closed without arrest" |
      disposition == "Open/No arrest")

```

### Create a Function to Return Odds Ratios (ORs) and CIs

```{r}

glm_sex_OR = function(glm){
  glm |> broom::tidy() |> 
  mutate(
    OR_estimate = exp(estimate),
    OR_CI_low = exp(estimate - 1.96 * std.error),
    OR_CI_high = exp(estimate + 1.96 * std.error)
    ) |>
  filter(term == "victim_sexMale") |>
  select(OR_estimate, OR_CI_low, OR_CI_high)
}

```

### Run a GLM for Baltimore Only

```{r}

homicides_baltimore = homicides |>
  filter(city_state == "Baltimore, MD") 

logit_model = glm(unsolved ~ victim_race + victim_age + victim_sex,
      data = homicides_baltimore, family = "binomial")

logit_model |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate),
    OR_CI_low = exp(estimate - 1.96 * std.error),
    OR_CI_high = exp(estimate + 1.96 * std.error)
    ) |>
  filter(term == "victim_sexMale") |>
  select(OR, OR_CI_low, OR_CI_high)

```

### Run a GLM for All Cities

```{r}

homicides_ORs = homicides |>
  nest(data = -city_state) |>
  mutate(
    models = map(data, \(df) glm(unsolved ~ victim_age + victim_race + victim_sex, 
                      data = df,
                      family = "binomial")),
    results = map(models, glm_sex_OR) 
    ) |> 
  select(city_state, results) |>
  unnest(results) 

homicides_ORs

```

### Make a Plot of Cities by OR

```{r}

homicides_ORs |>
  ggplot(aes(x = fct_reorder(city_state, OR_estimate), y = OR_estimate)) +
  geom_boxplot() +
  geom_errorbar(aes(ymin = OR_CI_low, ymax = OR_CI_high)) +
  coord_flip() +
  labs(
    title = "Odds Ratios between Males/Females for Unsolved Homicides Across the U.S.",
    x = "City",
    y = "Estimated Odds Ratio of Homicide Being Unsolved"
  ) 
    

```

From the plot, we can see that generally, homicides with male victims were more likely to be unsolved than those with female victims, since the odds ratios for most cities are greater than 1. The greatest estimated odds ratio, by far, is in New York City, where the estimate is nearly 4. That is, on average, male victims' homicides had nearly four times higher odds of being unsolved than female victims' homicides.

A few west-coast cities (Albuquerque, Stockton, and Fresno) actually had an odds ratio less than 1, implying that homicides with male victims were **less** likely to be unsolved than those with female victims.

## Problem 3

```{r}

birthweight = read_csv("data/birthweight.csv") |>
  mutate(
    ppbmi = round(ppbmi, 2)
  )

view(birthweight)
```

